### 2018-05-28 evaluation.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references

We have implemented SCP in the OpenBLAS~\cite{openblas} library (version 0.3.0-dev) and evaluated
it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of a cluster is almost the same as
the 4-core processor in Figure~\ref{fig:hierarchy}
except that the L2 caches of all 16 clusters connect to the main memory
with hardware coherence.
Table~\ref{tab:cluster} lists the
architectural parameters for the cores and caches
in this processor.
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers~\cite{top500},
relies on the DGEMM variant.

In our evaluation, we use the tiling factors
computed as in Section~\ref{subsec:example}:
$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$,
as this combination causes GEMM to deliver the best performance 
when SCP is not used.
The degree of parallelism is controlled by two parameters,
the number of active clusters $NC$
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \times NT_C$.
For the Phytium 2000+ processor, we have
$NC \in [1, 16]$ and $NT_C \in [1, 4]$. 
Generally, the DGEMM performance is measured in $flops$.
As $NT$ varies in our evaluation,
we also  use another metric derived from $flops$,
the average thread efficiency,
to describe our performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \times cdot \widehat{flops})$,
where $\widehat{flops}$ stands for the theoretical performance
peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

@@

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:drawback-4}
    \includegraphics[width=.31\textwidth]{figures/sec1-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:drawback-8}
    \includegraphics[width=.31\textwidth]{figures/sec1-8}
  }
  \\
  \subfigure[$NT=16$]{
    \label{fig:drawback-16}
    \includegraphics[width=.31\textwidth]{figures/sec1-16}
  }
  \subfigure[$NT=32,64$]{
    \label{fig:drawback-32}
    \includegraphics[width=.31\textwidth]{figures/sec1-32}
  }
  \caption{Average thread efficiency with cache sharing}
  \label{fig:drawback}
\end{figure}

@@

@@

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:benefit-4}
    \includegraphics[width=.31\textwidth]{figures/sec2-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:benefit-8}
    \includegraphics[width=.31\textwidth]{figures/sec2-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:benefit-16}
    \includegraphics[width=.31\textwidth]{figures/sec2-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:benefit-32}
    \includegraphics[width=.31\textwidth]{figures/sec2-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:benefit-64}
    \includegraphics[width=.31\textwidth]{figures/sec2-64}
  }
  \caption{Average thread efficiency under SCP}
  \label{fig:benefit}
\end{figure}

@@

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:papi-4}
    \includegraphics[width=.31\textwidth]{figures/papi-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:papi-8}
    \includegraphics[width=.31\textwidth]{figures/papi-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:papi-16}
    \includegraphics[width=.31\textwidth]{figures/papi-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:papi-32}
    \includegraphics[width=.31\textwidth]{figures/papi-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:papi-64}
    \includegraphics[width=.31\textwidth]{figures/papi-64}
  }
  \caption{Cache miss rates with $NT_C=4$}
  \label{fig:papi}
\end{figure}

@@

\begin{figure}
  \centering
  \subfigure[Conventional]{
    \label{fig:packb.conventional}
    \includegraphics[width=.50\textwidth]{figures/strategy-conventional}
  }
  \subfigure[Full privatization]{
    \label{fig:packb.full}
    \includegraphics[width=.50\textwidth]{figures/strategy-full}
  }
  \subfigure[Partial privatization]{
    \label{fig:packb.partial}
    \includegraphics[width=.50\textwidth]{figures/strategy-partial}
  }
  \caption{Packing workload ($B_2$) of thread $T_1$ with $NT_C=4$ and $NC=4$}
  %% \caption{Packing strategies of shared matrix $B_2$ with $NT_C=4$ and $NC=4$}
  \label{fig:packb}
\end{figure}

Privatization incurs more packing overhead because
the packing of $B_2$ must be done redundantly.
Figure~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelization configuration specified by
$NT_C=4$ and $NC=4$.
In each case, the entire packing workload of $B_2$ is divided into 16 tasks.
So only the packing workload in one thread, $T_1$, 
which is highlighted by the z-curves (as shown), is illustrated.
%% and the task assignment is annoted in Figure~\ref{fig:packb}.
In the conventional packing approach
(Figure~\ref{fig:packb.conventional}),
each thread takes a single task.
In full privatization
(Figure~\ref{fig:packb.full}),
all tasks are replicated to all
the threads.
As a result, the packing overhead grows in proportion to $NT$, which is unacceptable
if GEMM is highly parallelized, e.g., when $NT=64$.
In partial privatization 
(Figure~\ref{fig:packb.partial}), which is 
an alternative to full privatization,
privatization only occurs inside
a cluster so that the tasks for one cluster need not to be
replicated in the other clusters.
In this case, the extra packing overhead
is bounded by $NT_C$.
For example, on Phytium 2000+, the
overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in the conventional packing approach.

Figure~\ref{fig:privb} shows the results for partial
privatization with $NT_C=4$ and $NC=4$.
The results for the other 
configurations are similar.
For comparison purposes,
SCP with and without privatization,
denoted as SCP-P and SCP, respectively,
are given in Figure~\ref{fig:privb}.
Figure~\ref{fig:privb.papi} shows the L1 and L2
cache miss rates, which are 
slightly lower with privatization (SCP-P) than without
(SCP). Figure~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower L1 and L2 cache miss rates,
privatization (SCP-P) suffers a performance loss
(relative to SCP) due to 
the extra packing overhead introduced by privatization.
Figure~\ref{fig:privb.breakdown} compares the
three types of major overheads in percentage terms 
over the total execution time, including
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While behaving similarly as SCP
in (1) and (3), SCP-P is worse than SCP in (2).
To understand this further,
Table~\ref{tab:breakdown} gives a breakdown of each's
execution time in terms of (1) -- (3) and the
computation time of GEBP. As
the overhead for packing $B_2$
is multiplied by a factor of $NT_C=4$,
the total overhead has thus increased
from $2.35\%$ (under SCP) to $4.72\%$ (under SCP-P).

\begin{figure}
  \centering
  \subfigure[Cache miss rates]{
    \label{fig:privb.papi}
    \includegraphics[width=.31\textwidth]{figures/privb-papi}
  }
  \subfigure[Average thread efficiency]{
    \label{fig:privb.ate}
    \includegraphics[width=.31\textwidth]{figures/privb-ate}
  }
  \subfigure[Execution time breakdown]{
    \label{fig:privb.breakdown}
    \includegraphics[width=.65\textwidth]{figures/privb-breakdown}
  }
  \caption{Performance of partial privatization with $NT_C=4$ and $NC=4$}
  \label{fig:privb}
\end{figure}

@@

Our results demonstrate that privatizing the shared matrix $B_2$ can further
reduce inter-thread cache conflicts.
However, the improvement in cache performance is small.
As $B_2$ only occupies a small portion of the shared L2 cache,
this marginal improvement is not sufficient to
balance the extra packing overhead introduced
by privatization.

>>>>

We have implemented SCP in the OpenBLAS~\cite{openblas} library
%% (version 0.3.0-dev)
and evaluated it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of a cluster is almost the same as
the 4-core processor in Figure~\ref{fig:hierarchy}
except that the L2 caches of all 16 clusters connect to the main memory
with hardware coherence.
\textcolor{red}{
Table~\ref{tab:cluster} lists the
architectural parameters for the cores and caches
in this processor
and Table~\ref{tab:software} shows the software environment used in out evaluation.}
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers~\cite{top500},
relies on the DGEMM variant.

\begin{table}
  \centering
  \caption{Software environment used in evaluation}
  \label{tab:software}
  \begin{tabular}{lll}
    \toprule
    Environment & Software & Version \\
    \midrule
    Operating System & GNU/Linux & 4.4.0 AArch64 \\
    Compiler & GNU/GCC & 6.3.0 \\
    Thread Model & OpenMP & 4.0 \\
    BLAS & OpenBLAS & 0.3.0-dev \\
    \bottomrule
  \end{tabular}
\end{table}

In our evaluation, we use the tiling factors
computed as in Section~\ref{subsec:example}:
$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$,
as this combination causes GEMM to deliver the best performance 
when SCP is not used.
The degree of parallelism is controlled by two parameters,
the number of active clusters $NC$
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \times NT_C$.
For the Phytium 2000+ processor, we have
$NC \in [1, 16]$ and $NT_C \in [1, 4]$. 
Generally, the DGEMM performance is measured in $flops$.
As $NT$ varies in our evaluation,
we use another metric derived from $flops$,
the average thread efficiency,
to describe our performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \times \widehat{flops})$,
where $\widehat{flops}$ stands for the theoretical performance
peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

@@

@@

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:drawback-4}
    \includegraphics[width=.45\textwidth]{figures/sec1-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:drawback-8}
    \includegraphics[width=.45\textwidth]{figures/sec1-8}
  }
  \\
  \subfigure[$NT=16$]{
    \label{fig:drawback-16}
    \includegraphics[width=.45\textwidth]{figures/sec1-16}
  }
  \subfigure[$NT=32,64$]{
    \label{fig:drawback-32}
    \includegraphics[width=.45\textwidth]{figures/sec1-32}
  }
  \caption{Average thread efficiency with cache sharing}
  \label{fig:drawback}
\end{figure}

@@

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:benefit-4}
    \includegraphics[width=.45\textwidth]{figures/sec2-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:benefit-8}
    \includegraphics[width=.45\textwidth]{figures/sec2-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:benefit-16}
    \includegraphics[width=.45\textwidth]{figures/sec2-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:benefit-32}
    \includegraphics[width=.45\textwidth]{figures/sec2-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:benefit-64}
    \includegraphics[width=.45\textwidth]{figures/sec2-64}
  }
  \caption{Average thread efficiency under SCP}
  \label{fig:benefit}
\end{figure}

@@

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:papi-4}
    \includegraphics[width=.45\textwidth]{figures/papi-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:papi-8}
    \includegraphics[width=.45\textwidth]{figures/papi-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:papi-16}
    \includegraphics[width=.45\textwidth]{figures/papi-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:papi-32}
    \includegraphics[width=.45\textwidth]{figures/papi-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:papi-64}
    \includegraphics[width=.45\textwidth]{figures/papi-64}
  }
  \caption{Cache miss rates with $NT_C=4$}
  \label{fig:papi}
\end{figure}

@@

\begin{figure}
  \centering
  \subfigure[Conventional]{
    \label{fig:packb.conventional}
    \includegraphics[width=.65\textwidth]{figures/strategy-conventional}
  }
  \subfigure[Full privatization]{
    \label{fig:packb.full}
    \includegraphics[width=.65\textwidth]{figures/strategy-full}
  }
  \subfigure[Partial privatization]{
    \label{fig:packb.partial}
    \includegraphics[width=.65\textwidth]{figures/strategy-partial}
  }
  \caption{Packing workload ($B_2$) of thread $T_1$ with $NT_C=4$ and $NC=4$}
  %% \caption{Packing strategies of shared matrix $B_2$ with $NT_C=4$ and $NC=4$}
  \label{fig:packb}
\end{figure}

Privatization incurs more packing overhead because
the packing of $B_2$ must be done redundantly.
Figure~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelization configuration specified by
$NT_C=4$ and $NC=4$.
In each case, the entire packing workload of $B_2$ is divided into 16 tasks.
So only the packing workload of one thread, $T_1$, 
which is highlighted by the z-curves (as shown), is illustrated.
%% and the task assignment is annotated in Figure~\ref{fig:packb}.
In the conventional packing approach
(Figure~\ref{fig:packb.conventional}),
each thread takes a single task.
In full privatization
(Figure~\ref{fig:packb.full}),
all tasks are replicated to all
the threads.
As a result, the packing overhead grows in proportion to $NT$, which is unacceptable
if GEMM is highly parallelized, e.g., when $NT=64$.
In partial privatization 
(Figure~\ref{fig:packb.partial}), which is 
an alternative to full privatization,
privatization only occurs inside
a cluster so that the tasks for one cluster need not to be
replicated in the other clusters.
In this case, the extra packing overhead
is bounded by $NT_C$.
For example, on Phytium 2000+, the
overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in the conventional packing approach.

Figure~\ref{fig:privb} shows the results for partial
privatization with $NT_C=4$ and $NC=4$.
The results for the other 
configurations are similar.
For comparison purposes,
SCP with and without privatization,
denoted as SCP-P and SCP, respectively,
are both given in Figure~\ref{fig:privb}.
Figure~\ref{fig:privb.papi} shows the L1 and L2
cache miss rates, which are 
slightly lower with privatization (SCP-P) than without
(SCP). Figure~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower L1 and L2 cache miss rates,
privatization (SCP-P) suffers a performance loss
(relative to SCP) due to 
the extra packing overhead introduced by privatization.
Figure~\ref{fig:privb.breakdown} compares the
three types of major overheads in percentage terms 
over the total execution time, including
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
\textcolor{red}{
The solid and dashed horizontal lines represent
the average value of total overhead across matrix sizes
under SCP and SCP-P, respectively.}
While behaving similarly as SCP
in (1) and (3), SCP-P is worse than SCP in (2).
To understand this further,
Table~\ref{tab:breakdown} gives a breakdown of each's
execution time in terms of (1) -- (3) and the
computation time of GEBP. As
the overhead for packing $B_2$
is multiplied by a factor of $NT_C=4$,
the total overhead has thus increased
from $2.35\%$ (under SCP) to $4.72\%$ (under SCP-P).

\begin{figure}
  \centering
  \subfigure[Cache miss rates]{
    \label{fig:privb.papi}
    \includegraphics[width=.45\textwidth]{figures/privb-papi}
  }
  \subfigure[Average thread efficiency]{
    \label{fig:privb.ate}
    \includegraphics[width=.45\textwidth]{figures/privb-ate}
  }
  \subfigure[Execution time breakdown]{
    \label{fig:privb.breakdown}
    \includegraphics[width=.90\textwidth]{figures/privb-breakdown}
  }
  \caption{Performance of partial privatization with $NT_C=4$ and $NC=4$}
  \label{fig:privb}
\end{figure}

@@

Our results demonstrate that privatizing the shared matrix $B_2$ can further
reduce inter-thread cache conflicts.
However, the improvement in cache performance is small.
As $B_2$ only occupies a small portion of the shared L2 cache during GEMM execution,
this marginal improvement is not sufficient to
balance the extra packing overhead introduced
by privatization.

### 2018-05-28 related.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references

>>>>

\textcolor{red}{
Because GEMM is highly optimized and parallelized,
other higher level dense linear algebra operations,
such as Cholesky, LU and QR factorization,
can just simply call GEMM to benefit from multi-core and many-core processors.
In other words, this programming pattern in dense linear algebra software
harness a fork-and-join style parallelism.
To reduce synchronization overhead,
The PLASMA project~\cite{plasma2009,plasma2010,plasma2017}
utilizes a task-driven parallelization paradigm
in which dense linear algorithms are formulated
into directed-acyclic-graphs (DAG) of fine-grained tasks.
The tasks are dynamically scheduled based on data dependencies
between them (represented by edges in the DAG).
The principle of SCP also applies to this DAG based parallelization.
Input matrices of different tasks can be packed into
different sets of a shared cache and
tasks can be scheduled based on the cache sets their input matrices reside in,
i.e., tasks whose input matrices live in the same cache sets
are scheduled to the same core in priority.
}

### 2018-05-28 scp.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references

Given the tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
SCP will systematically determine 
the memory layout for the packed matrices $A_2$ and $B_2$ in GEMM.
The memory layout of the other packed matrices at
lower layers are determined once the memory layout of
$A_2$ and $B_2$ is determined.
We assume that the share matrix $B_2$ is packed
in the conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

We make a few standard assumptions, which usually
hold for the architecture considered:
\begin{itemize}
\item The architecture has 2 or 3 levels of caches; %% i.e. $NL=2$ or $NL=3$
\item All caches are inclusive set-associative caches; %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
\item The caches at the same level are homogeneous; and %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
\item A cache's way-capacity is a multiple of the sum of way-capacities of its children,
i.e., $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
\end{itemize}

@@

In Algorithm~\ref{alg:scp.phase1} (Phase 1),
the per-thread sizes for memory buffers $\mathcal{D}_p$ and
$\mathcal{D}_s$ are computed
(lines~\ref{line:size.p} -- \ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Then $align$ is found
(lines~\ref{line:align.init} -- \ref{line:align.endfor}) 
and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
Specifically, $align$ is initialized with the L1 cache line size $l_1$.
Then for each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
the earlier $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight on using $wc_l/nt_l$ is that
each way of the level-$l$ cache should be
equally partitioned among the $nt_l$ threads
in set-partitioning.
Finally, memory is allocated (lines~\ref{line:alloc.begin} -- \ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin} -- \ref{line:d.end}).

>>>>

Given the tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
SCP will systematically determine 
the memory layout for the packed matrices $A_2$ and $B_2$ in GEMM.
The memory layout of the other packed matrices at
lower layers are determined once the memory layout of
$A_2$ and $B_2$ is determined.
We assume that the shared matrix $B_2$ is packed
in the conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

We make a few standard assumptions, which usually
hold for the architecture considered:
\begin{itemize}
\item The architecture has 2 or 3 levels of caches. %% i.e. $NL=2$ or $NL=3$
\item All caches are inclusive set-associative caches. %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
\item The caches at the same level are homogeneous. %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
\item A cache's way-capacity is a multiple of the sum of way-capacities of its children,
i.e., $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
\end{itemize}

@@

In Algorithm~\ref{alg:scp.phase1} (Phase 1),
the per-thread sizes for memory buffers $\mathcal{D}_p$ and
$\mathcal{D}_s$ are computed
(lines~\ref{line:size.p} -- \ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Then $align$ is found
(lines~\ref{line:align.init} -- \ref{line:align.endfor}) 
and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
Specifically, $align$ is initialized with the L1 cache line size $l_1$.
Then for each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
the earlier $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight on using $wc_l/nt_l$ is that
each way of the level-$l$ cache should be
equally partitioned among the $nt_l$ threads
in set-partitioning.
Finally, memory is allocated (lines~\ref{line:alloc.begin} -- \ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (lines~\ref{line:d.begin} -- \ref{line:d.end}).
